### 2018-09-18 pilots/pilot3/vacc_manyBelief_pilot3.Rmd

---
title: "Vaccines: Many Beliefs Pilot 3"
author: "Derek Powell"
date: "June 26, 2017"
output: 
  html_notebook: 
    code_folding: hide
---

```{r}

filename <- "data/vb_pilot3.csv"

headers <- as.matrix(read.csv(filename, skip = 0, header = F, nrows = 1, as.is = T ))
df <- read.csv(filename, skip = 3, header = F)
colnames(df) <- headers

df <- df[which(df$Status==0),] # remove survey previews

# remove unused qualtrics variables
remove_cols <- c("RecipientLastName",
                 "RecipientFirstName",
                 "RecipientEmail",
                 "Finished",
                 "ResponseId",
                 "ExternalReference",
                 "DistributionChannel",
                 "UserLanguage",
                 "Status")

df <- df[ , -which(names(df) %in% remove_cols)]
```

```{r}
all_items <- c("hb_hb_1",
"hb_hb_2",
"hb_hb_3",
"hb_hb_4",
"nat_nat_1",
"nat_nat_2r",
"nat_nat_3",
"nat_nat_4",
"nat_nat_5",
"medSkept_medSkept_1",
"medSkept_medSkept_2r",
"medSkept_medSkept_3",
"medSkept_medSkept_4",
"medSkept_medSkept_5",
"medSkept_medSkept_6",
"diseaseRare_diseaseRare_1",
"diseaseRare_diseaseRare_2",
"diseaseRare_diseaseRare_3",
"diseaseRare_diseaseRare_4",
"diseaseSevere_diseaseSevere_1",
"diseaseSevere_diseaseSevere_2",
"diseaseSevere_diseaseSevere_3",
"diseaseSevere_diseaseSevere_4r",
"diseaseSevere_diseaseSevere_5",
"diseaseSevere_diseaseSevere_6",
"vaccEff_vaccEff_1",
"vaccEff_vaccEff_2r",
"vaccEff_vaccEff_3r",
"vaccEff_vaccEff_4",
"vaccDanger_vaccDanger_1",
"vaccDanger_vaccDanger_2",
"vaccDanger_vaccDanger_3",
"vaccDanger_vaccDanger_4r",
"vaccTox_vaccTox_1",
"vaccTox_vaccTox_2",
"vaccTox_vaccTox_3",
"vaccTox_vaccTox_4r",
"vaccStrain_vaccStrain_1",
"vaccStrain_vaccStrain_2r",
"vaccStrain_vaccStrain_3",
"vaccStrain_vaccStrain_4",
"vaccStrain_vaccStrain_5",
"vaccIntent_vaccIntent_1r",
"vaccIntent_vaccIntent_2",
"vaccIntent_vaccIntent_3r",
"vaccIntent_vaccIntent_4",
"vaccIntent_vaccIntent_5r",
"parentExpert_parentExpert_1",
"parentExpert_parentExpert_2",
"parentExpert_parentExpert_3",
"parentExpert_parentExpert_4r",
"infantImmLimCap_infantImmLimCap_1",
"infantImmLimCap_infantImmLimCap_2r",
"infantImmLimCap_infantImmLimCap_3",
"infantImmLimCap_infantImmLimCap_4",
"infantImmLimCap_infantImmLimCap_5",
"immSympt_immSympt_1",
"immSympt_immSympt_2r",
"immSympt_immSympt_3r",
"immSympt_immSympt_4",
"immSympt_immSympt_5r",
"immSympt_immSympt_6",
"immSympt_immSympt_7r",
"infantImmWeak_infantImmWeak_1",
"infantImmWeak_infantImmWeak_2",
"infantImmWeak_infantImmWeak_3",
"infantImmWeak_infantImmWeak_4r",
"overpar_overpar_1r",
"overpar_overpar_2r",
"overpar_overpar_3",
"overpar_overpar_4",
"overpar_overpar_5")

```

```{r}
library(ggplot2)
# df <- read.csv("vb_pilot2.csv")
df <- df[which(df$diseaseRare_check_5==5),]

reverse_items <- all_items[!sapply(
  lapply(all_items, function (x) {
  if (substr(x,nchar(x),nchar(x))=="r") {
    return(x)
  }
}), is.null)]

df[,reverse_items] <- lapply(df[,reverse_items], function(x) {
  8-x
})

df_main <- df[,all_items]
```

# Introduction

This is a notebook for the first pilot experiment of our (Kara, Ellen, and Derek) first pilot experiment examining vaccine theories using multiple attitude measures. We recruited 100 participants and lost 6 to comprehension checks, leaving 94.

# Item Correlations

Here's a correlation matrix for all individual items. It appears items of some scales are intercorrelating well, though others are weaker. There is also considerable correlation among items across scales--as we might expect.

```{r,fig.height=9,fig.width=9}
library(corrplot)
library(viridis)

corrplot(cor(df_main),
         tl.cex=.65,
         tl.col="black",
         method="color",
         col=viridis(128))
```

Here's another matrix, this time sorted according to a hierarchical clustering algorithm, with bounding rectangles identifying clusters. Whereas we treated these as 15 different scales, this finds just 3 separate clusters. I don't think this is a problem--we expected and even hoped these scales would be correlated with one another.

```{r,fig.height=9,fig.width=9}
corrplot(cor(df_main[,all_items]),
         order="hclust",
         # addrect=3,
         tl.cex=.65,
         tl.col="black",
         method="color",
         col=viridis(128))
```

# Scale Internal Consistency Reliability

Below are computed chronbach's alpha for all scales along with more detailed analyses of scales that are problematic.

## Holistic Balance

```{r}
hb <- c("hb_hb_1",
"hb_hb_2",
"hb_hb_3",
"hb_hb_4")

hb.alpha <- psych::alpha(df[,hb])
hb.alpha$total$raw_alpha
```

## Naturalism

```{r}
nat <- c("nat_nat_1",
"nat_nat_2r",
"nat_nat_3",
"nat_nat_4",
"nat_nat_5")

naturalism.alpha <- psych::alpha(df[,nat])
naturalism.alpha$total$raw_alpha

corrplot(cor(df[,nat]),
         order="alpha",
         tl.cex=.65,
         tl.col="black",
         method="color",
         addCoef.col="black",
         col=viridis(128))

```

## Disease Rarity

**NOTE**: we previously scored this one kind of strangely, so it's now been relabeld as "diseaseRare" to indicate that it reflects participants' belief the diseases are extremely rare.

```{r}
diseaseRare <- c("diseaseRare_diseaseRare_1",
"diseaseRare_diseaseRare_2",
"diseaseRare_diseaseRare_3",
"diseaseRare_diseaseRare_4")

diseaseRare.alpha <- psych::alpha(df[,diseaseRare])
diseaseRare.alpha$total$raw_alpha

```

## Disease Severity

Disease severe is now a highly reliable scale (and could even remove an item.)

```{r}
diseaseSevere <- c("diseaseSevere_diseaseSevere_1",
"diseaseSevere_diseaseSevere_2",
"diseaseSevere_diseaseSevere_3",
"diseaseSevere_diseaseSevere_4r",
"diseaseSevere_diseaseSevere_5",
"diseaseSevere_diseaseSevere_6")

diseaseSevere.alpha <- psych::alpha(df[,diseaseSevere])
diseaseSevere.alpha$total$raw_alpha
diseaseSevere.alpha

```
### Diagnosing

Here are histograms of each item

```{r}
library(ggplot2)
library(reshape2)

ggplot(data = melt(df[,diseaseSevere]), mapping = aes(x = value)) + 
    geom_histogram(bins = 7) + facet_wrap(~variable, scales = 'free_x') + xlim(1,7)
corrplot(cor(df[,diseaseSevere]),
         order="alpha",
         tl.cex=.65,
         tl.col="black",
         method="color",
         addCoef.col="black",
         col=viridis(128))

```

## Vaccine Effectiveness

```{r}
vaccEff <- c("vaccEff_vaccEff_1",
"vaccEff_vaccEff_2r",
"vaccEff_vaccEff_3r",
"vaccEff_vaccEff_4")

vaccEff.alpha <- psych::alpha(df[,vaccEff])
vaccEff.alpha$total$raw_alpha
```

## Vaccine Danger

```{r}
vaccDanger <- c("vaccDanger_vaccDanger_1",
"vaccDanger_vaccDanger_2",
"vaccDanger_vaccDanger_3",
"vaccDanger_vaccDanger_4r")

vaccDanger.alpha <- psych::alpha(df[,vaccDanger])
vaccDanger.alpha$total$raw_alpha
```

## Vaccines are toxic

```{r}
vaccTox <- c("vaccTox_vaccTox_1",
"vaccTox_vaccTox_2",
"vaccTox_vaccTox_3",
"vaccTox_vaccTox_4r")

vaccTox.alpha <- psych::alpha(df[,vaccTox])
vaccTox.alpha$total$raw
```

## Medical Skepticism

```{r}
medSkept <- c("medSkept_medSkept_1",
"medSkept_medSkept_2r",
#"medSkept_medSkept_3", # planned to remove
"medSkept_medSkept_4",
"medSkept_medSkept_5",
"medSkept_medSkept_6")

medSkept.alpha <- psych::alpha(df[,medSkept])
medSkept.alpha$total$raw_alpha
medSkept.alpha
```

## Vaccination Intentions

```{r}
vaccIntent <- c("vaccIntent_vaccIntent_1r",
"vaccIntent_vaccIntent_2",
"vaccIntent_vaccIntent_3r",
"vaccIntent_vaccIntent_4",
"vaccIntent_vaccIntent_5r")

vaccIntent.alpha <- psych::alpha(df[,vaccIntent])
vaccIntent.alpha$total$raw_alpha
```

## Parental Expertise

Success!

```{r}
parExpert <- c("parentExpert_parentExpert_1",
"parentExpert_parentExpert_2",
"parentExpert_parentExpert_3",
"parentExpert_parentExpert_4r")

parExpert.alpha <- psych::alpha(df[,parExpert])
parExpert.alpha$total$raw_alpha
parExpert.alpha
```
```{r}
library(ggplot2)
library(reshape2)

ggplot(data = melt(df[,parExpert]), mapping = aes(x = value)) + 
    geom_histogram(bins = 7) + facet_wrap(~variable, scales = 'free_x') + xlim(1,7)

corrplot(cor(df[,parExpert]),
         order="AOE",
         tl.cex=.65,
         tl.col="black",
         method="color",
         addCoef.col="black",
         col=viridis(128))

```

## Infant immune system has limited capacity

Reliability is still weak Might improve if we ditch my new item #5

```{r}
immCap <- c(
  "infantImmLimCap_infantImmLimCap_1",
"infantImmLimCap_infantImmLimCap_2r",
"infantImmLimCap_infantImmLimCap_3",
"infantImmLimCap_infantImmLimCap_4",
"infantImmLimCap_infantImmLimCap_5")

immCap.alpha <- psych::alpha(df[,immCap])
immCap.alpha$total$raw_alpha
immCap.alpha

```

```{r}

library(ggplot2)
library(reshape2)

ggplot(data = melt(df[,immCap]), mapping = aes(x = value)) + 
    geom_histogram(bins = 7) + facet_wrap(~variable, scales = 'free_x') + xlim(1,7)

corrplot(cor(df[,immCap]),
         order="AOE",
         tl.cex=.65,
         tl.col="black",
         method="color",
         addCoef.col="black",
         col=viridis(128))
```

## Immune system produces cold symptoms

Reliability actually got worse!

```{r}
immSympt <- c("immSympt_immSympt_1",
"immSympt_immSympt_2r",
"immSympt_immSympt_3r",
"immSympt_immSympt_4",
"immSympt_immSympt_5r",
"immSympt_immSympt_6",
"immSympt_immSympt_7r")

immSympt.alpha <- psych::alpha(df[,immSympt])
immSympt.alpha$total$raw_alpha
immSympt.alpha

```

No obvious fixes here -- just a mess.

```{r}
# library(ggplot2)
# library(reshape2)
# 
# ggplot(data = melt(df[,immSympt]), mapping = aes(x = value)) + 
#     geom_histogram(bins = 7) + facet_wrap(~variable, scales = 'free_x') + xlim(1,7)
# 
# corrplot(cor(df[,immSympt]),
#          order="AOE",
#          tl.cex=.65,
#          tl.col="black",
#          method="color",
#          addCoef.col="black",
#          col=viridis(128))
```

```{r}
library(ggplot2)
library(reshape2)

# ggplot(data = melt(df[,immSympt]), mapping = aes(x = value)) + 
#     geom_histogram(bins = 7) + facet_wrap(~variable, scales = 'free_x') + xlim(1,7)

corrplot(cor(df[,c(immCap,immWeak)]),
         order="AOE",
         tl.cex=.5,
         tl.col="black",
         method="color",
         addCoef.col="black",
         col=viridis(128))
```

## Infants' immune systems are weak

Looks pretty good.

```{r}
immWeak <- c("infantImmWeak_infantImmWeak_1",
"infantImmWeak_infantImmWeak_2",
"infantImmWeak_infantImmWeak_3",
"infantImmWeak_infantImmWeak_4r")

immWeak.alpha <- psych::alpha(df[,immWeak])
immWeak.alpha$total$raw_alpha
immWeak.alpha

```

## Vaccines strain the immune system

Now with five items this makes for a very reliable scale.

```{r}
vaccStrain <- c("vaccStrain_vaccStrain_1",
"vaccStrain_vaccStrain_2r",
"vaccStrain_vaccStrain_3",
"vaccStrain_vaccStrain_4",
"vaccStrain_vaccStrain_5")

# cor(df$immStrain_1,df$immStrain_2r)

vaccStrain.alpha <- psych::alpha(df[,vaccStrain])

vaccStrain.alpha

```

## Overparenting with babies (new scale)

This is a brand new scale so these numbers are ok but probably want this improved. Could be improved slightly by removing item #2.

```{r}
overPar <- c("overpar_overpar_1r",
"overpar_overpar_2r",
"overpar_overpar_3",
"overpar_overpar_4",
"overpar_overpar_5")

# cor(df$immStrain_1,df$immStrain_2r)

overPar.alpha <- psych::alpha(df[,overPar])

overPar.alpha
```

```{r}
ggplot(data = melt(df[,overPar]), mapping = aes(x = value)) + 
    geom_histogram(bins = 7) + facet_wrap(~variable, scales = 'free_x') + xlim(1,7)

corrplot(cor(df[,overPar]),
         order="AOE",
         tl.cex=.65,
         tl.col="black",
         method="color",
         addCoef.col="black",
         col=viridis(128))
```

# Scale Distributions

Here I'm computing the scale average for all scales but the "children's weak immune system" scale and the "passive immune system" scale, where I'm making the changes suggested from the reliability analysis. *UPDATE*: Now also changing the "immune system and cold symptoms" scale to remove a bad item.

```{r}
df_scales<-df_main
df_scales$holBalance <- rowMeans(df[,hb])
df_scales$naturalism <- rowMeans(df[,nat])
df_scales$diseaseRare <- rowMeans(df[,diseaseRare])
df_scales$diseaseSevere <- rowMeans(df[,diseaseSevere])
df_scales$vaccEff <- rowMeans(df[,vaccEff])
df_scales$vaccDanger <- rowMeans(df[,vaccDanger])
df_scales$medSkept <- rowMeans(df[,medSkept])
df_scales$vaccIntent <- rowMeans(df[,vaccIntent])
# df_scales$parDuty <- rowMeans(df[,parDuty])
df_scales$parExpert <- rowMeans(df[,parExpert])
df_scales$immCap <- rowMeans(df[,immCap])
df_scales$immSympt <- rowMeans(df[,immSympt])
df_scales$immWeak <- rowMeans(df[,immWeak])
df_scales$vaccStrain <- rowMeans(df[,vaccStrain])
df_scales$vaccTox <- rowMeans(df[,vaccTox])
df_scales$overParenting <- rowMeans(df[,overPar])
```

```{r,fig.height=9,fig.width=9}
library(ggplot2)
library(reshape2)

scale_vars <- c("holBalance",
                "naturalism",
                "diseaseRare",
                "diseaseSevere",
                "vaccEff",
                "vaccDanger",
                "medSkept",
                "vaccIntent", 
                "parExpert",
                "immCap",
                "immSympt",
                "immWeak",
                "vaccStrain",
                "vaccTox",
                "overParenting")

# df_hist <- df_scales[,scale_vars]
plt.hist <- ggplot(data = melt(df_scales[,scale_vars]), mapping = aes(x = value)) + 
    geom_histogram(bins = 14) + facet_wrap(~variable, scales = 'free_x') + xlim(1,7)

# ggsave("scale_hists.png",plt.hist)

plt.hist
```

# Scale correlations

Here are intercorrelations among the scales (ordered by angular order of eigenvectors). I think we should use a big grain of salt interpreting any of these because this is all based on just `r nrow(df_scales)` paricipants' data. Nevertheless, let's take a look.

Some of this is expected and some is a bit surprising. Of course once we start taking into account other conditional dependencies, some of these correlations could change, or even reverse sign.

```{r,fig.height=9,fig.width=9}
corrplot(cor(df_scales[,scale_vars]),
         order="hclust",
         tl.cex=1.25,
         tl.col="black",
         method="color",
         addCoef.col="black",
         col=viridis(128))

```

```{r,fig.height=9,fig.width=9}
corrplot(cor(df_scales)[73:87,1:72],
         tl.cex=.65,
         tl.col="black",
         method="color",
         col=viridis(128))
```

```{r}
df_long <- df_scales
df_long$y_val <- df_long$vaccIntent
df_long <- reshape(df_long, idvar="workerId", timevar = "item", direction="long", 
             varying=list(x_val = scale_vars[-8]),
             v.names = c("x_val"))
df_long$item <- factor(df_long$item,
                       levels = seq(1,14,1),
                       labels=scale_vars[-8])

plt.scatter <- ggplot(data = df_long, mapping = aes(x = x_val, y=y_val)) + 
    geom_point(shape=1) + 
  geom_smooth(method="lm", se=FALSE, col=viridis_pal()(5)[3]) + 
  facet_wrap(~item) + 
  xlim(1,7) + ylim(1,7) + labs(y="Vaccine Intent", x="")

ggsave("scale_scatter.png",plt.scatter)
```