### 2018-09-18 study1/vacc_manyBelief_study1.Rmd

---
title: 'Vaccines: Many Beliefs Study 1'
author: "Derek Powell, Kara Weisman"
date: "September 7, 2017"
output: html_notebook
---

```{r setup}
library(tidyverse)
library(stringr)
library(corrplot)
library(viridis)
library(psych)
library(glmpath)
library(bnlearn)
```

```{r}
# moved all data import/munging to a separate file 
# so we can have multiple workbooks working withthe same data

source("vacc_import_data.R", chdir = TRUE)
```

# Introduction

This is a notebook for the first experiment of our (Kara, Ellen, and Derek) study examining vaccine theories using multiple attitude measures. We recruited 1200 and are left with `r length(d_demo$ResponseId)` after excluding any participants who failed at least 1 of the 2 attention checks.

# Demographics

A quick look at demographics variables...

## Study duration

```{r duration}
ggplot(d_demo, aes(x = Duration/60)) +
  geom_histogram() +
  geom_vline(xintercept = median(d_demo$Duration/60), color = "blue") +
  annotate("label", x = 0, y = 350, color = "blue", hjust = 0,
           label = paste0("median: ", round(median(d_demo$Duration/60)), "min")) +
  labs(x = "duration (min)") +
  theme_bw()
```

## Gender

```{r sex}
ggplot(d_demo, aes(x = sex)) +
  geom_bar() +
  labs(x = "gender") +
  theme_bw() +
  coord_flip()
```

## Age

```{r age}
ggplot(d_demo, aes(x = age)) +
  geom_histogram() +
  geom_vline(xintercept = median(d_demo$age), color = "blue") +
  annotate("label", x = 0, y = 150, color = "blue", hjust = 0,
           label = paste0("median: ", round(median(d_demo$age)), "y")) +
  labs(x = "age (y)") +
  theme_bw()
```

## Race

```{r race}
ggplot(d_demo, aes(x = race)) +
  geom_bar() +
  labs(x = "race") +
  theme_bw() +
  coord_flip()
```

## Religion

```{r religion}
ggplot(d_demo, aes(x = religion)) +
  geom_bar() +
  labs(x = "religion") +
  theme_bw() +
  coord_flip()
```

## Education

```{r education}
ggplot(d_demo, aes(x = educ)) +
  geom_bar() +
  labs(x = "educ") +
  theme_bw() +
  coord_flip()
```

## Income

```{r income}
ggplot(d_demo, aes(x = income)) +
  geom_bar() +
  labs(x = "income") +
  theme_bw() +
  coord_flip()
```

## Parent

```{r parent}
ggplot(d_demo, aes(x = parent)) +
  geom_bar() +
  labs(x = "parent") +
  theme_bw() +
  coord_flip()
```

## Expecting

```{r expecting}
ggplot(d_demo, aes(x = expecting)) +
  geom_bar() +
  labs(x = "expecting") +
  theme_bw() +
  coord_flip()
```

## Number of children

```{r children}
ggplot(d_demo, aes(x = factor(children))) +
  geom_bar() +
  labs(x = "number of children") +
  theme_bw() +
  coord_flip()
```

## Age of youngest child

```{r youngest_child}
ggplot(d_demo, aes(x = factor(youngest_child))) +
  geom_bar() +
  labs(x = "age of youngest child") +
  theme_bw() +
  coord_flip()
```

## Comments on study experience

```{r comments}
d_demo %>% select(comments)
```

# Heatmap

```{r heatmap, fig.width = 5, fig.height = 5}
# make dataframe for heatmpa, efa, etc.
d_efa <- d_wide %>% select(-c(StartDate, Duration, sex:comments)) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId")

# make correlation plot heatmap
# corrplot(cor(d_efa),
#          tl.cex=.75,
#          tl.col="black",
#          method="color",
#          col=viridis(128))

# with clustering
corrplot(cor(d_efa),
         order="hclust",
         # addrect=3,
         tl.cex=.75,
         tl.col="black",
         method="color",
         col=viridis(128))
```

# Exploratory factor analysis

Just for fun, let's do exploratory factor analysis!

```{r efa}
# NEED TO AUTOMATE THIS: do maximal unrotated efa, retain factors with SS loadings > 1 and Prop Explained > 0.05
fa_maximal <- fa(d_efa, nfactors = floor(length(levels(factor(d_long$question)))/3), rotate = "none")
nfactors_efa <- 4

# small rotated efa (oblimin rotation, doesn't have to be orthogonal)
fa_small <- fa(d_efa, nfactors = nfactors_efa, rotate = "oblimin")
fa.sort(fa_small$loadings)
```

By my typical factor retention criteria, we observe for 4 factors here:

1. Beliefs/attitudes about vaccines and medical skepticism
2. Beliefs/attitudes about infants
3. Beliefs/attitudes about holistic balance and naturalism
4. Beliefs/attitudes about childhood diseases

```{r parallel analysis}
# extract number of factors recommended by fa.parallel
fa_parallel <- fa(d_efa, nfactors = fa.parallel(d_efa, plot = F)$nfact, rotate = "oblimin")
fa.sort(fa_parallel$loadings)
```

By anoother criterion (parallel analysis) we observe 12ish factors. For the most part, these correspond to the 14 intended scales, with a little bit of combination in the factors related to vaccines per se:

1. Vaccine intentions + some other beliefs/attitudes about vaccines (*vaccIntent* + *vaccEff* + some of *vaccDanger*)
2. Disease severity (*diseaseSevere*)
3. Vaccines strain the immune system + some other beliefs/attitudes about vaccines (*vaccStrain* + some of *vaccDanger*)
4. Infant immune system is weak (*infantImmWeak*)
5. Infant immune system has limited capacity (*infantImmLimCap*)
6. Disease rarity (*diseaseRare*)
7. Holistic balance (*hb*)
8. Parent expertise (*parentExpert*)
9. Overparenting (*overpar*)
10. Vaccines are toxic (*vaccTox* + some of *vaccDanger*)
11. Naturalism (*nat*)
12. Medical skepticism (*medSkept*)

Now let's leave the vaccIntent scale out of this (holding it out as our DV) and do traditional-style EFA for all the rest of the items:

```{r efa without vaccIntent}
# NEED TO AUTOMATE THIS: do maximal unrotated efa, retain factors with SS loadings > 1 and Prop Explained > 0.05
fa_maximal2 <- fa(d_efa %>% select(-starts_with("vaccIntent")), 
                 nfactors = floor(length(levels(factor(d_long$question)))/3), 
                 rotate = "none")
nfactors_efa2 <- 4

# small rotated efa (oblimin rotation, doesn't have to be orthogonal)
fa_small2 <- fa(d_efa %>% select(-starts_with("vaccIntent")), 
                                 nfactors = nfactors_efa2, 
                                 rotate = "oblimin")
fa.sort(fa_small2$loadings)
```

By my typical factor retention criteria, we again observe for 4 factors here - very similar to the 4 factors we observed when we left the vaccIntent scale in:

1. Beliefs/attitudes about vaccines, medical skepticism, parental expertise, and naturalism
2. Beliefs/attitudes about infants' immune systems (and a little overparenting)
3. Beliefs/attitudes about childhood diseases
4. Beliefs/attitudes about holistic balance and overparenting (and a little naturalism)

# Scale reliability analyses

As a reminder, our goal was $\alpha$ >= 0.80 for each scale individually.

## Alpha for each scale

### Disease rarity (*diseaseRare*)

```{r reliability diseaseRare}
alpha_diseaseRare <- d_long %>%
  filter(question_block == "diseaseRare") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_diseaseRare
```

For *diseaseRare*, $\alpha$ = `r round(alpha_diseaseRare$total$raw_alpha, 2)`. This is *just* under our cutoff (ok).

### Disease severity (*diseaseSevere*)

```{r reliability diseaseSevere}
alpha_diseaseSevere <- d_long %>%
  filter(question_block == "diseaseSevere") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_diseaseSevere
```

For *diseaseSevere*, $\alpha$ = `r round(alpha_diseaseSevere$total$raw_alpha, 2)`. This is well over our cutoff (good).

### Holistic balance (*hb*)

```{r reliability hb}
alpha_hb <- d_long %>%
  filter(question_block == "hb") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_hb
```

For *hb*, $\alpha$ = `r round(alpha_hb$total$raw_alpha, 2)`. This is a little over our cutoff (good).

### Infant immune system has limited capacity (*infantImmLimCap*)

```{r reliability infantImmLimCap}
alpha_infantImmLimCap <- d_long %>%
  filter(question_block == "infantImmLimCap") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_infantImmLimCap
```

For *infantImmLimCap*, $\alpha$ = `r round(alpha_infantImmLimCap$total$raw_alpha, 2)`. This is well over our cutoff (good).

### Infant immune system is weak (*infantImmWeak*)

```{r reliability infantImmWeak}
alpha_infantImmWeak <- d_long %>%
  filter(question_block == "infantImmWeak") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_infantImmWeak
```

For *infantImmWeak*, $\alpha$ = `r round(alpha_infantImmWeak$total$raw_alpha, 2)`. This is well over our cutoff (good).

### Medical skepticism (*medSkept*)

```{r reliability medSkept}
alpha_medSkept <- d_long %>%
  filter(question_block == "medSkept") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_medSkept
```

For *medSkept*, $\alpha$ = `r round(alpha_medSkept$total$raw_alpha, 2)`. This is just under our cutoff (ok).

### Naturalism (*nat*)

```{r reliability nat}
alpha_nat <- d_long %>%
  filter(question_block == "nat") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_nat
```

For *nat*, $\alpha$ = `r round(alpha_nat$total$raw_alpha, 2)`. This is just under our cutoff (ok).

### Overparenting (*overpar*)

```{r reliability overpar}
alpha_overpar <- d_long %>%
  filter(question_block == "overpar") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_overpar
```

For *overpar*, $\alpha$ = `r round(alpha_overpar$total$raw_alpha, 2)`. This is quite a bit under our cutoff (not great).

### Parental expertise (*parentExpert*)

```{r reliability parentExpert}
alpha_parentExpert <- d_long %>%
  filter(question_block == "parentExpert") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_parentExpert
```

For *parentExpert*, $\alpha$ = `r round(alpha_parentExpert$total$raw_alpha, 2)`. This is just over our cutoff (good).

### Vaccines are dangerous (*vaccDanger*)

```{r reliability vaccDanger}
alpha_vaccDanger <- d_long %>%
  filter(question_block == "vaccDanger") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_vaccDanger
```

For *vaccDanger*, $\alpha$ = `r round(alpha_vaccDanger$total$raw_alpha, 2)`. This is well over our cutoff (good).

### Vaccines are effective (*vaccEff*)

```{r reliability vaccEff}
alpha_vaccEff <- d_long %>%
  filter(question_block == "vaccEff") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_vaccEff
```

For *vaccEff*, $\alpha$ = `r round(alpha_vaccEff$total$raw_alpha, 2)`. This is well over our cutoff (good).

### Vaccine intentions (*vaccIntent*)

```{r reliability vaccIntent}
alpha_vaccIntent <- d_long %>%
  filter(question_block == "vaccIntent") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_vaccIntent
```

For *vaccIntent*, $\alpha$ = `r round(alpha_vaccIntent$total$raw_alpha, 2)`. This is just over our cutoff (good).

### Vaccines strain the immune system (*vaccStrain*)

```{r reliability vaccStrain}
alpha_vaccStrain <- d_long %>%
  filter(question_block == "vaccStrain") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_vaccStrain
```

For *vaccStrain*, $\alpha$ = `r round(alpha_vaccStrain$total$raw_alpha, 2)`. This is well over our cutoff (good).

### Vaccines are toxic (*vaccTox*)

```{r reliability vaccTox}
alpha_vaccTox <- d_long %>%
  filter(question_block == "vaccTox") %>%
  select(ResponseId, question, response_num_rev) %>%
  spread(question, response_num_rev) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  alpha()

alpha_vaccTox
```

For *vaccTox*, $\alpha$ = `r round(alpha_vaccTox$total$raw_alpha, 2)`. This is well over our cutoff (good).

## Scale reliability summary

Things worked out pretty well:

- 10 scales were above the $\alpha$ = 0.80 cutoff (*diseaseSevere*, *hb*,*infantImmLimCap*, *infantImmWeak*, *parentExpert*, *vaccDanger*, *vaccEff*, *vaccIntent*, *vaccStrain*, *vaccTox*)
- 3 scales were just under the cutoff (*diseaseRare*: $\alpha$ = `r round(alpha_diseaseRare$total$raw_alpha, 2)`; *medSkept*: $\alpha$ = `r round(alpha_medSkept$total$raw_alpha, 2)`; *nat*: $\alpha$ = `r round(alpha_nat$total$raw_alpha, 2)`)
- 1 scale was pretty well below the cutoff (*overpar*: $\alpha$ = `r round(alpha_overpar$total$raw_alpha, 2)`)

So, we should probably be a little cautious in interpreting results related to overparenting (*overpar*) (though note that even that reliability is not terrible relative to what other research groups use!).

# Scale score distributions

Let's look at the distributions of scores on each scale. I scored each scale by taking the average of the participant's responses to items on that scale - so for each scale, scores could theoretically range from -3 (participant said "Strongly disagree" to all items after reverse-coding) to +3 (participant said "Strongly agree" to all items after reverse-coding).

```{r score distributions, fig.width = 5, fig.asp = 0.67}
score_medians <- d_sum %>%
  mutate(scale = recode_factor(
    question_block,
    "vaccIntent" = "Vaccination Intentions",
    "nat" = "Naturalism",
    "hb" = "Holistic Balance",
    "overpar" = "Protective Parenting",
    "parentExpert" = "Parental Expertise",
    "medSkept" = "Medical Skepticism",
    "vaccDanger" = "Vaccine Danger",
    "vaccTox" = "Toxic Additives in Vaccines",
    "vaccEff" = "Vaccine Efficacy",
    "diseaseRare" = "Disease Rarity",
    "diseaseSevere" = "Disease Severity",
    "infantImmWeak" = "Infant Immune System:\nWeakness",
    "infantImmLimCap" = "Infant Immune System:\nLimited Capacity",
    "vaccStrain" = "Infant Immune System:\nVaccines Strain")) %>%
  group_by(scale) %>%
  summarise(median_score = median(score))

d_sum %>%
  mutate(scale = recode_factor(
    question_block,
    "vaccIntent" = "Vaccination Intentions",
    "nat" = "Naturalism",
    "hb" = "Holistic Balance",
    "overpar" = "Protective Parenting",
    "parentExpert" = "Parental Expertise",
    "medSkept" = "Medical Skepticism",
    "vaccDanger" = "Vaccine Danger",
    "vaccTox" = "Toxic Additives in Vaccines",
    "vaccEff" = "Vaccine Efficacy",
    "diseaseRare" = "Disease Rarity",
    "diseaseSevere" = "Disease Severity",
    "infantImmWeak" = "Infant Immune System:\nWeakness",
    "infantImmLimCap" = "Infant Immune System:\nLimited Capacity",
    "vaccStrain" = "Infant Immune System:\nVaccines Strain")) %>%
  ggplot(aes(x = score)) +
  facet_wrap(~ scale, ncol = 5) +
  geom_histogram(binwidth = 0.25) +
  scale_x_continuous(limits = c(-3, 3), breaks = seq(-3, 3, 1)) +
  geom_vline(data = score_medians, aes(xintercept = median_score), 
             color = "blue") +
  geom_label(data = score_medians,
             aes(x = -3, y = 200, 
                 label = paste0("median: ", round(median_score, 2))),
             color = "blue", hjust = 0) +
  labs(x = "Score", y = "Number of participants") +
  theme_bw()
```

Most of these scales are pretty normally distributed, with the major exceptions of *diseaseSevere* and *vaccIntent*, which are both severely skewed, and a few other more minor ceiling/floor issues (e.g., *vaccDanger*, *vaccEff*, *vaccTox*).

# First-order correlations

```{r}
# with clustering
corrplot(cor(d_bn),
         order="hclust",
         # addrect=3,
         tl.cex=.75,
         tl.col="black",
         method="color",
         col=viridis(128))
```

```{r, fig.width = 4, fig.asp = 0.8}
temp_order <- corrMatOrder(cor(d_bn))

temp_cor <- cor(d_bn)[temp_order, temp_order] %>%
  data.frame() %>%
  rownames_to_column("scale1") %>%
  mutate(order1 = 1:14) %>%
  gather(scale2, cor, -c(scale1, order1))

temp_cor_df <- temp_cor %>%
  full_join(temp_cor %>% 
              distinct(scale1, order1) %>%
              rename(scale2 = scale1, order2 = order1)) %>%
  mutate_at(vars(scale1, scale2),
            funs(recode(.,
                        "diseaseRare" = "Disease Rarity",
                        "diseaseSevere" = "Disease Severity",
                        "hb" = "Holistic Balance",
                        "infantImmLimCap" = "Infant Immune System: Limited Capacity",
                        "infantImmWeak" = "Infant Immune System: Weakness",
                        "medSkept" = "Medical Skepticism",
                        "nat" = "Naturalism",
                        "overpar" = "Protective Parenting",
                        "parentExpert" = "Parental Expertise",
                        "vaccDanger" = "Vaccine Danger",
                        "vaccEff" = "Vaccine Efficacy",
                        "vaccIntent" = "Vaccination Intentions",
                        "vaccStrain" = "Infant Immune System: Vaccines Strain",
                        "vaccTox" = "Toxic Additives in Vaccines")))

ggplot(temp_cor_df,
       aes(x = reorder(scale1, desc(order1)), 
           y = reorder(scale2, order2), 
           fill = cor)) +
  geom_tile(color = "black") +
  geom_text(aes(label = format(round(cor, 2), digits = 2)), size = 3) +
  scale_fill_viridis("R", limits = c(-1, 1),
                     guide = guide_colorbar(barheight = 15)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        axis.title = element_blank())
```

# Regularized regression

Here's another thing I know how to do, so why not try it...

```{r glmpath}
# make a matrix for predictors
predictors <- d_sum %>% 
  select(ResponseId, question_block, score) %>%
  arrange(ResponseId, question_block) %>%
  filter(question_block != "vaccIntent") %>%
  spread(question_block, score) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  as.matrix()

# make a matrix for outcome
outcome <- d_sum %>%
  select(ResponseId, question_block, score) %>%
  arrange(ResponseId, question_block) %>%
  filter(question_block == "vaccIntent") %>%
  spread(question_block, score) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId") %>%
  as.matrix()

glm_fit <- glmpath(predictors, outcome, family = "gaussian"); glm_fit
# plot(glm_fit, xvar = "lambda")

d_glm <- glm_fit$b.predictor %>% # extract the coefficient estimates
  data.frame() %>% # turn them into a dataframe
  rownames_to_column(var = "step") %>% # add the rownames as a "step" variable
  mutate(lambda = glm_fit$lambda, # add the lambda values for each step
         aic = glm_fit$aic, # add AIC for each step
         bic = glm_fit$bic) %>% # add BIC for each step
  gather(scale, coeff, diseaseRare:vaccTox) %>% # turn into longform
  arrange(step, scale) # arrange by step number and scale

d_glm_labs <- d_glm %>% filter(lambda == 0) %>% select(scale, coeff)

ggplot(data = d_glm, aes(x = lambda, y = coeff, color = scale)) +
  geom_point() +
  geom_line() + 
  scale_x_reverse(limits = c(1200, -300)) +
  geom_text(data = d_glm_labs, aes(x = -10, y = coeff, label = scale), hjust = 0) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Order of predictors based on regularized regression",
       subtitle = "Most important predictors' coefficients become non-zero first (when lambda is large)")
```

Regularized regression suggests that, in the aggregate, beliefs that vaccines are *not* dangerous (opposite of *vaccDanger*) and that vaccines are effective (*vaccEff*) are particularly strong predictors of intentions to vaccinate.

The next most important predictors is a belief that vaccines are *not* toxic (opposite of *vaccTox*), followed by a belief that do *not* strain the immune system (opposite of *vaccStrain*), a belief that childhood diseases are severe (*diseaseSevere*), and a *lack* of medical skepticism (opposite of *medSkept*).

Other predictors seem less important, according to this analysis.

**Important**: This does *not* take into account all the causal relationships among the different scales that we are primarily interested in - it just examines each scale and its direct relationship to our outcome variable (*vaccIntent*). (Perhaps these will end up being the most proximal causal relationships...?)

# Bayesian network structure learning

A first attempt!

Resources used:

- [package guide](https://cran.r-project.org/web/packages/bnlearn/bnlearn.pdf)
- [hameddaily.blogspot.com](http://hameddaily.blogspot.com/2015/02/bayesian-network-in-r-introduction.html)
- [UC Denver Biostatistics dept.](http://www.ucdenver.edu/academics/colleges/PublicHealth/Academics/departments/Biostatistics/WorkingGroups/Documents/Networks%20Presentation%20With%20Sachs%20-%20032317.pdf)

```{r bn dataframe}
d_bn <- d_sum %>%
  select(ResponseId, question_block, score) %>%
  spread(question_block, score) %>%
  remove_rownames() %>%
  column_to_rownames("ResponseId")
```

## Learning the structure

There are lots of different methods... still figuring out what they mean.

### Method 1: Grow-shrink

From the package info: "Based on the Grow-Shrink Markov Blanket, the first (and simplest) Markov
blanket detection algorithm used in a structure learning algorithm."

```{r bn grow-shrink, fig.width = 6, fig.height = 6}
res_bn_gs <- gs(d_bn)
plot(res_bn_gs)
```

### Method 2: Incremental association

From the package info: "based on the Markov blanket detection algorithm of the same name, which is based on a two-phase selection scheme (a forward selection followed by an attempt to remove false positives)."

```{r bn iamb, fig.width = 6, fig.height = 6}
res_bn_iamb <- iamb(d_bn)
plot(res_bn_iamb)
```

### Method 3: Fast incremental association

From the package info: "a variant of IAMB which uses speculative stepwise forward selection to reduce the number of conditional independence tests."

```{r bn fast.iamb, fig.width = 6, fig.height = 6}
res_bn_fast.iamb <- fast.iamb(d_bn)
plot(res_bn_fast.iamb)
```

### Method 4: Interleaved incremental association

From the package info: "another variant of IAMB which uses forward stepwise selection to avoid false positives in the Markov blanket detection phase."

```{r bn inter.iamb, fig.width = 6, fig.height = 6}
res_bn_inter.iamb <- inter.iamb(d_bn)
plot(res_bn_inter.iamb)
```

### Method 5: Hill-climbing

From the package info: "a hill climbing greedy search on the space of the directed graphs. The optimized implementation uses score caching, score decomposability and score equivalence to reduce the number of duplicated tests."

```{r bn hc, fig.width = 6, fig.height = 6}
res_bn_hc <- hc(d_bn)
plot(res_bn_hc)
```

### Method 6: Tabu search

From the package info: "a modified hill-climbing able to escape local optima by selecting a network
that minimally decreases the score function."

```{r bn tabu, fig.width = 6, fig.height = 6}
res_bn_tabu <- tabu(d_bn)
plot(res_bn_tabu)
```

### Method 7: Max-min hill-climbing

From the package info: "a hybrid algorithm which combines the Max-Min Parents and Children algorithm (to restrict the search space) and the Hill-Climbing algorithm (to find the optimal network structure in the restricted space)."

```{r bn mmhc, fig.width = 6, fig.height = 6}
res_bn_mmhc <- mmhc(d_bn)
plot(res_bn_mmhc)
```

### Method 8: Restricted maximization

From the package info: "a more general implementation of the Max-Min Hill-Climbing, which can use any combination of constraint-based and score-based algorithms."

```{r bn rsmax2, fig.width = 6, fig.height = 6}
res_bn_rsmax2 <- rsmax2(d_bn)
plot(res_bn_rsmax2)
```

Let's go with MMHC (Method #7) for now (for no real reason other than the fact that I've heard of it before).

## Parameter learning

```{r bn parameter learning}
res_bn_mmhc_fitted <- bn.fit(res_bn_mmhc, data = d_bn); res_bn_mmhc_fitted
```

## Limit to just parents?

```{r bn just parents, fig.width = 6, fig.height = 6}
d_bn_parents <- d_bn %>%
  rownames_to_column("ResponseId") %>%
  left_join(d_demo %>% select(ResponseId, parent)) %>%
  filter(parent == "Yes") %>%
  select(-parent) %>%
  column_to_rownames("ResponseId")

res_bn_parents_mmhc <- mmhc(d_bn_parents)
plot(res_bn_parents_mmhc)
```