### 2019-08-08 notebooks/cross-study-analysis.Rmd

---
title: "R Notebook"
output: html_notebook
---

The relevant [NY Times front page article](https://www.nytimes.com/2019/04/29/health/measles-outbreak-cdc.html).

```{r}
# # load s1 data (original cogsci dataset)
source("../scripts/load-data-s1.R")

s1 <- d_sum %>%
  rename(scale = question_block) %>%
  group_by(scale, workerId) %>% 
  summarize(s1 = mean(score, na.rm = T)) %>%
  ungroup()

# load s2 data (disease risk intervention)
source("../scripts/load-data-s2.R")

s2 <- d_scored %>%
  # rename(subid = workerId) %>%
  filter(phase == "pre") %>% 
  group_by(scale, workerId) %>% 
  summarize(s2 = mean(mean, na.rm = T)) %>%
  ungroup()

# load S3 data (vaccine toxin intervention)

source("../scripts/load-data-s3.R")

s3 <- df_scored %>%
  # rename(subid = workerId) %>%
  group_by(scale, workerId) %>% 
  summarize(s3 = mean(pre, na.rm = T) - 4) %>% # put on -3, 3 scale
  ungroup()

source("../scripts/load-data-s4.R")

library(lubridate)

s4 <- df4 %>%
  mutate(prior_study = case_when(
    workerId %in% s1$workerId ~ "S1",
    workerId %in% s3$workerId ~ "S3",
    TRUE ~ "None"
  )) %>%
  mutate(resp = resp-4) %>%
  filter(prior_study != "S3", 
         StartDate < ymd_hms("2019-05-19:00:00:00")) %>% # exclude Ps from s3, some of whom saw intervention
  # filter(prior_study == "None") %>% # Look only at fresh Ps
  group_by(scale, workerId) %>%
  summarize(s4 = mean(resp))

s5 <- df4 %>%
  mutate(resp = resp-4) %>%
  filter(StartDate > ymd_hms("2019-05-19:00:00:00")) %>% # exclude Ps from s3, some of whom saw intervention
  # filter(prior_study == "None") %>% # Look only at fresh Ps
  group_by(scale, workerId) %>%
  summarize(s5 = mean(resp))
```

```{r}
# merge dataframes

df_studies <-  s1 %>%
  full_join(s2) %>%
  full_join(s3) %>%
  full_join(s4) %>%
  full_join(s5) %>%
  gather(study, score, -c(workerId, scale)) %>%
  filter(!is.na(study), !is.na(score)) %>%
  group_by(scale) %>%
  mutate(score_std = scale(score, scale = T)) %>%
  ungroup()
```

## Changes across studies

```{r}
df_studies %>%
  group_by(study, scale) %>%
  summarize(
    mean = mean(score),
    se = sd(score)/sqrt(n()),
    ul = mean + se,
    ll = mean - se,
    ul95 = mean + 1.96*se,
    ll95 = mean - 1.96*se
  ) %>%
  ungroup() %>%
  mutate(study = ordered(study, 
                         labels=c("S1\n(9/'17)", 
                                  "S2\n(5/'18)", 
                                  "S3\n(4/'19)", 
                                  "S4\n(5/01)", 
                                  "S5\n(5/20)")
                         )
         ) %>%
  ggplot(aes(x=study, y = mean, ymin =ll, ymax=ul)) +
  geom_pointrange() +
  facet_wrap(~scale, scales="free") +
  theme_bw(base_size = 12) +
  labs(y="Mean Response", x = "Study")
```

```{r}
# important to recognize these are small changes in absolute terms ...
df_studies %>%
  ggplot(aes(x=study, y = score)) +
  geom_violin(fill = "blue",color="white",alpha=.5, scale="width", draw_quantiles=c(.5)) +
  facet_wrap(~scale, scales="free") +
  theme_bw()
```

Model changes across studies: I'm going to make a beta regression model with study as the predictor

```{r}
library(betareg)
library(brms)

rescale_beta <- function(x, lower, upper) {
  # rescales onto the open interval (0,1)
  # rescales over theoretical bounds of measurement, specified by "upper" and "lower"

  N <- length(x)
  res <- (x - lower) / (upper - lower)
  res <- (res * (N - 1) + .5) / N

  return(as.vector(res))
}

fit <- betareg(score ~ study,
               data = df_studies %>%
                 filter(scale =="vaccIntent") %>%
                 mutate(score = rescale_beta(score,-3,3))
               )

summary(fit)

# fit_brm <- brm(
#   score ~ study,
#   data = df_studies %>%
#                  filter(scale =="vaccIntent") %>%
#                  mutate(score = rescale_beta(score,-3,3)),
#   family = Beta("logit"),
#   prior = prior(normal(0,1), class="b"),
#   iter = 2000,
#   warmup = 1000,
#   chains = 1,
#   cores = 1
# )

```

## Within-subjects test

let's focus on the clearest test of whether and how opinions have changed, a within-subjects comparison among our re-recruited participants. These are participants who completed all 14 scales over 18 months ago, and are now returning and completing them all again.

```{r}
combined <- df4 %>%
  filter(workerId %in% s1$workerId) %>%
  select(workerId, scale, item, resp) %>%
  group_by(workerId,scale) %>%
  summarize(posttest = mean(resp)-4) %>%
  right_join(s1) %>%
  rename(pretest = s1) %>%
  mutate(returned = !is.na(posttest))

combined %>%
  filter(returned == TRUE) %>%
  ggplot(aes(x=pretest, y=posttest)) +
  geom_jitter(alpha=.5) +
  facet_wrap(~scale) +
  theme_bw()

combined_long <- combined %>%
  gather(phase, resp, pretest, posttest)

combined_long %>%
  group_by(returned, phase, scale) %>%
  summarize(
    mean = mean(resp),
    se = sd(resp)/sqrt(n()),
    ul = mean + se,
    ll = mean - se,
    ul95 = mean + 1.96*se,
    ll95 = mean - 1.96*se
  ) %>%
  ungroup() %>%
  mutate(phase = ordered(phase, levels=c("pretest","posttest"))) %>%
  ggplot(aes(x=phase, y = mean, ymin =ll, ymax=ul, color=returned)) +
  geom_point(position=position_dodge(width=.25), size=2, shape=16) +
  geom_errorbar(position=position_dodge(width=.25), size=.75, width=0, alpha=.8) +
  geom_errorbar(aes(ymin=ll95, ymax = ul95), width=0, position=position_dodge(width=.25), alpha=.5) +
  facet_wrap(~scale, scales="free") +
  theme_bw(base_size=18)

combined %>%
  filter(returned==TRUE) %>%
  mutate(diff = posttest-pretest) %>%
  group_by(scale) %>%
  summarize(
    mean = mean(diff),
    se = sd(diff)/sqrt(n()),
    ul = mean + se,
    ll = mean - se,
    ul95 = mean + 1.96*se,
    ll95 = mean - 1.96*se
  ) %>%
  ungroup() %>%
  # mutate(phase = ordered(phase, levels=c("pretest","posttest"))) %>%
  ggplot(aes(x=reorder(scale, mean), y = mean, ymin =ll, ymax=ul)) +
  geom_point(position=position_dodge(width=.25), size=2, shape=16) +
  geom_errorbar(position=position_dodge(width=.25), size=.75, width=0, alpha=.8) +
  geom_errorbar(aes(ymin=ll95, ymax = ul95), width=0, position=position_dodge(width=.25), alpha=.5) +
  geom_hline(yintercept = 0, linetype="dashed",alpha=.5) +
  # facet_wrap(~scale, scales="free") +
  theme_bw(base_size = 18) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(y = "Change scores", x = "Belief scale")
```

```{r}
diff_df <- combined %>% 
  group_by(scale) %>% 
  summarize(posttest = mean(posttest, na.rm=TRUE), 
            pretest=mean(pretest)) %>% 
  mutate(diff = posttest-pretest)

diff_df
```

Have vaccine intentions increased within subjects from 18 months ago? YES.

We could test this with a t-test, but I think it might be better to use a somewhat more sophisticated model. I'm going to test for differences across phases using beta regression. (I'm using "pretest" and "posttest" even though that's not qutie accurate, there was no intervention)

```{r}
t.test(resp ~ phase, data = combined_long %>% filter(returned, scale=="vaccIntent"), paired=TRUE)
```

```{r}

library(betareg)

fit <- betareg(resp ~ phase, data = combined_long %>% 
                 mutate(resp = rescale_beta(resp, -3, 3)) %>%
                 filter(scale=="vaccEff")
               )

summary(fit)
```

Do changes relate to media consumpion? (keeping in mind that's a noisy measure ...)

Yes, for our binary yes/no response variable, but no for our news frequency scale (Could try monotonic effects -- not seeing it at all vs seeing it is a much bigger difference than seeing it once vs twice, etc)

```{r}

df_prepost_news <- df4 %>%
  filter(workerId %in% s1$workerId) %>%
  select(workerId, scale, item, resp, 
         vacc_news, vacc_news_freq, news_favorable, outbreaks, vacc_conv
         ) %>%
  group_by(workerId,scale,vacc_news, vacc_news_freq, news_favorable, outbreaks, vacc_conv ) %>%
  summarize(posttest = mean(resp)-4) %>%
  left_join(s1) %>%
  rename(pretest = s1) %>%
  mutate(returned = !is.na(posttest)) %>%
  ungroup() %>%
  mutate(vacc_news_freq = ifelse(vacc_news==0, 0.0, vacc_news_freq))

fit <- betareg(
  posttest ~ pretest + vacc_news,
  data = df_prepost_news %>%
  filter(scale =="vaccIntent") %>%
  mutate(posttest = rescale_beta(posttest,-3,3))
)

summary(fit)
```

## Participants who saw news

```{r}
df_prepost_news %>%
  filter(returned==TRUE, !is.na(vacc_news)) %>% # somehow 5? participants slippd by without answering 
  mutate(diff = posttest-pretest) %>%
  mutate(vacc_news = as.factor(vacc_news)) %>%
  group_by(vacc_news,scale) %>%
  summarize(
    mean = mean(diff),
    se = sd(diff)/sqrt(n()),
    ul = mean + se,
    ll = mean - se,
    ul95 = mean + 1.96*se,
    ll95 = mean - 1.96*se
  ) %>%
  ungroup() %>%
  mutate(vacc_news = factor(vacc_news, c(1,0), labels=c("Yes","No"))) %>%
  # mutate(phase = ordered(phase, levels=c("pretest","posttest"))) %>%
  ggplot(aes(x=reorder(scale, mean), y = mean, ymin =ll, ymax=ul, color=vacc_news)) +
  geom_point(position=position_dodge(width=.25), size=2, shape=16) +
  geom_errorbar(position=position_dodge(width=.25), size=.75, width=0, alpha=.8) +
  geom_errorbar(aes(ymin=ll95, ymax = ul95), width=0, position=position_dodge(width=.25), alpha=.5) +
  geom_hline(yintercept = 0, linetype="dashed",alpha=.5) +
  # facet_wrap(~scale, scales="free") +
  theme_bw(base_size = 18) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "right") +
  labs(y = "Change scores", x = "Belief scale", color ="News\nexposure")

```

Are the changes coherent? What is driving the changes?

Let's see how they correlate imagining all changes are driven by different beliefs (these are approximations of what model will predict)

```{r message=F}
cors <- cor(d_bn)

# for (s in colnames(d_bn)){
#   
#   cors_df <- tibble(scale = names(cors[s,]), w = cors[s,])
#   
#   z <- cors_df %>% 
#   left_join(diff_df) %>%
#   mutate(pred_diff = w * diff_df[diff_df$scale==s,"diff"][[1]]) 
#   
#   print( paste(s,":",cor(z$pred_diff, z$diff)) )
# }

check_cor <- function(s){
  cors_df <- tibble(scale = names(cors[s,]), w = cors[s,])

  z <- cors_df %>% 
  left_join(diff_df) %>%
  mutate(pred_diff = w * diff_df[diff_df$scale==s,"diff"][[1]])

  return(cor(z$pred_diff, z$diff))
}

change_cors <- tibble(scale = colnames(d_bn), corr = map_dbl(scale, check_cor))

change_cors %>%
  ggplot(aes(y=reorder(scale, corr), x = corr)) +
  geom_point() +
  theme_bw(base_size=18) +
  theme(aspect.ratio=1) +
  labs(y="Scale", x = "Correlation")
```

Now, imagine it's a direct increase in intention to vaccinate ...

```{r}

targetted_belief <- "vaccIntent"

cors_df <- tibble(scale = names(cors[targetted_belief,]), w = cors[targetted_belief,])

z <- cors_df %>% 
left_join(diff_df) %>%
mutate(pred_diff = w * diff_df[diff_df$scale==targetted_belief,"diff"][[1]])

z %>%
  ggplot(aes(x=pred_diff, y=diff)) +
  geom_point() +
  theme_bw(base_size=18) +
  theme(aspect.ratio=1) +
  labs(x="Predicted change",y="Observed change", title=targetted_belief)

cor(z$pred_diff, z$diff)
```

That's a strong correlation!

# insepcting other ps

For some reason our "fresh" participants don't look quite like our S1 and S3 participants. A difference between S3 participants seems plausible, but I don't see how they could meaningfully differ from S1 participants beyond by chance. I say that especially because returning and non-returning participants from S1 didn't look different across any of the scales.

```{r}
df4 %>%
  mutate(prior_study = case_when(
    workerId %in% s1$workerId ~ "S1",
    workerId %in% s3$workerId ~ "S3",
    TRUE ~ "None"
  )) %>%
  mutate(resp = resp-4) %>%
  # filter(prior_study != "S3") %>% # exclude Ps from s3, some of whom saw intervention
  group_by(prior_study,scale, workerId) %>%
  summarize(s4 = mean(resp)) %>%
  group_by(prior_study, scale) %>%
  summarize(Mean = mean(s4, na.rm=TRUE)) %>%
  spread(prior_study, Mean)
```